---
title: "Reproducible and scalable data analysis workflows with targets"
subtitle: "Dynamic Function-Oriented Make-Like Declarative Pipelines for R"
author: 
  - .b[Matthew T. Warkentin], MSc. Ph.D (c)
institute: "Lunenfeld-Tanenbaum Research Institute, .b[.sinai-blue[Sinai] .sinai-orange[Health] .sinai-red[System]]"
date: 'Novemer 12, 2020'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: 
      - css/xaringan-theme.css
      - css/style.css
    includes:
      in_header: head.html
    seal: false
    nature:
      titleSlideClass: ["bottom", "left"]
      slideNumberFormat: "%current%"
      highlightStyle: atom-one-light
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
      navigation:
        scroll: false
---

name: title
class: left middle hide-count title-bg

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source("xaringan-theme.R")
source("packages.R")
source("setup.R")$value
```

```{r title-slide, echo=FALSE, message = FALSE, warning = FALSE}
htmltools::withTags(
  div(
    class = "talk-meta",
    div(
      h1(rmarkdown::metadata$title, class = "talk-title"),
      p(rmarkdown::metadata$subtitle, class = "talk-subtitle")
    ),
    div(
      class = "talk-author", 
      span(fontawesome::fa('id-badge', fill = '#FFFFFF'), rmarkdown::metadata$author),
      div(
        class = "talk-social pl3 pt2 pb2 lh-copy",
        paste0(
          div(fontawesome::fa('envelope', fill = '#FFFFFF'),
            a(href = "mailto:warkentin@lunenfeld.ca",
              "warkentin@lunenfeld.ca")),
          div(fontawesome::fa('github', fill = '#FFFFFF'),
            a(href = "https://github.com/mattwarkentin",
              "@mattwarkentin")),
          div(fontawesome::fa('twitter', fill = '#FFFFFF'),
            a(href = "https://twitter.com/mattwrkntn",
              "@mattwrkntn")),
          collapse = "<br/>"
          )
        ),
      span(
        fontawesome::fa('home', fill = '#FFFFFF'),
        rmarkdown::metadata$institute
        )
    ),
    div(
      class = "talk-date",
      span(
        fontawesome::fa('calendar-alt', fill = '#FFFFFF'),
        rmarkdown::metadata$date
      )
    )
  )
)
```

---

class: lh-copy

## What is `{targets}`?

.pull-left[
- `{targets}` is a project workflow tool that is very `R`-centric
  + Similar tools exist for other languages, such as `{GNU make}` and `{snakemake}`
  + `{drake}` was the predecessor to `{targets}` from from same author

- It allows you to effectively modularize your data analysis projects to create obvious and reproducible workflows

- Can easily extend your workflow to massively parallelize tasks
  + With some setup can use external compute resources (_e.g._ HPC) as a computational back-end for your pipeline
]

```{r echo=FALSE}
htmltools::img(src = "imgs/targets.png", id = "targets-pic")
```

---

What about `{drake}`?

---

## Why should I use `{targets}`?

- .dark-blue[__Organization__]
  + Explicitly building your projects as an end-to-end workflow can help keep your project more organized and focused (_i.e._ less interactive exploration)

--

- .dark-blue[__Modularity__]
  + Breaking your tasks into small digestible code chunks makes it easier to debug your code and easier to see how all of the parts fit together

--

- .dark-blue[__Transparency and Reproducibility__]
  + Out of the box, you get a transparent data analysis workflow which is easily reproducible 

--

- .dark-blue[__Caching and History__]
  + Re-running code that doesn't change is tedious and time-consuming. Caching results means you only run whats absolutely necessary to get up-to-date-results

--

- .dark-blue[__Scalability and Parallel Computing__]
  + Mental models of projects break down at scale. Building projects as workflows scales well and easily facilitates parallel computing


---

## `{targets}` etymology

---


https://wlandau.github.io/targets/
